Oriented Gradient Histogram applied to P300 Detection

### 2017-12-11 response.letter.tex

2.1) the title is misleading since a first, and naive, read suggest being able to classify and characterize any EEG with their different components (i.e., P200, P300, P600, N100, N200, N400 to name a few). Clearly, the title does not match the content since the authors focus on the P300

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. The title of the article was changed accordingly to your suggestion.  
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

2.2) there is evidence of the benefits of working with these signals P300, but, is the proposed method extensible to other signals?, this is not clear from the manuscript.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
You're  right, thank you very much for your comment.  We published in Ramele et al 2016 the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to decode Motor Imagery.  We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and to decode SSVEP patterns. Modificar%We modified the article to emphasize more clearly that the presented algorithm is  applicable to P300.

@@

% Como haces para que las referencias que estan aca en la respuesta apunten al articulo ???
%An explanation  of which is a satisfactory result for Algorithm~\ref{Algoritmo} was added in sections~\ref{errorevaluation} and~\ref{sec:EdgeSimulatedData}.  
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex
\subsection*{\ovalbox{Third Paragraph}}

3.1) Another issue with the manuscript is its mathematical vagueness. For example, the variable T is used as a vector or two-tuple but its notation is different from the notation the authors use for vectors

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation.  We worked hard to improve all the mathematical description of the method. Section FEATURE was modified as well as section PARAMETERS.  Section SIGNAL PLOT was revised and its mathematical terms were corrected.}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

3.2) Moreover, there is something wrong in equation (4)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
We appreciate a lot your truthful comment.  Equation (4) is very important for us because it is the essence of the method.  %We based that equation on the description of how the histogram of oriented gradients is calculated published in Vedaldi and Szeliski, Computer Vision and Algorithms.  
We reviewed the equation, clarified their terms and we hope it is now more clear.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

3.3) and something inconsistent between equations (7) and (8)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
We have revised the manuscript and corrected equations (7) and (8) which specify the ration between the signal properties and the SIFT's geometric parameters. 
We thank you for noticing this.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

3.4) Also, the authors go back and forth the digital space and the continuous space and although there is an equivalence, the manuscript could benefit from a clearer explanation.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation.  We modified all equations to emphasize the point that they are calculated in digital space. We also added that image gradients are also calculated using finite differences.
With this, we hope we have addressed your point.
 }
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

@@

4.1) Another issue that I find worrisome is the design of experiments.\\

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. 
We fully addressed your point.
 }
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

4.2) it seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your suggestion. The original dataset was published at the BNCI-Horizon website, which is very popular in BCI research.  Our approach emphasize the structure of the signal in terms of their morphological components in their visual description.% THIS IS TOUGH TO JUSTIFY 
In order to address this issue, we included a new dataset that we generated ourselves, on healthy subjects that we hope now improves the number of different subjects and that can help to validate the method.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

### 2017-12-11 article.tex
cambios de hoy

On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered. A local image patch is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm. It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center.

@@

\noindent  where $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $, $ i,j \in \{1,2,3,4\} $, $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al. in~\citep{Lowe2004,Vedaldi2010}.  Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.  As the patch is of size $4 \times 4$ and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained. It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $ , so the method considers both, the magnitude and the orientation of the gradient vector.

@@

\begin{enumerate}
\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
\item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively. The parameter $k_a$ is an input of the algorithm, and corresponds to the number of repetitions.
\item Compute  
\begin{equation}
x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c)
\label{averaging}
\end{equation}

@@

Parameters are selected according to the experimental protocol. First of all, the patch size in pixels is $S_P = 3s.4 \times 3s.4$. The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $.  Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event. With this purpose in consideration, the $s$ value election is essential.
%necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
We propose the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
\begin{equation}
s_x = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
\label{eq:mapping2}
\end{equation}

@@

\noindent where $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and $ \lambda $ is the length in seconds covered by the patch. The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.  This also provides a resolution of $ 1 $ Pixel representing $  \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{Fs \cdot \gamma}$ seconds in the horizontal direction. Finally, descriptors  $\mathbf{kp}$  are located at $ x =  0.55 Fs.\gamma = 35 $ and $ y = z(c) $ (Eq.~\ref{eq:zerolevel}).   In this way the whole transient event should be captured.

% esta figura no está comentada en el texto.
\begin{figure}[H]
\centering
\includegraphics[width=10cm]{patchgeometry.pdf}
\caption{The local patch captures the signal information based on both the vertical and horizontal scale.}
\label{fig:sampledescriptor2}
\end{figure}

>>>>

On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered. A local image patch of size $S_p \times S_p$ is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm. It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, then $S_p = 4.3s $ pixels.

@@

\noindent  where $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $, $ i,j \in \{1,2,3,4\} $, $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al. in~\citep{Lowe2004,Vedaldi2010}.  Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.  As the patch has  $16$ blocks and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained. It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $ , so the method considers both, the magnitude and the orientation of the gradient vector.

@@

\begin{enumerate}
\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
\item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively. The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
\item Compute  
\begin{equation}
x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c)
\label{averaging}
\end{equation}

@@

%Parameters are selected according to the experimental protocol. 
The patch size is $S_P = 3s.4 \times 3s.4$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm. The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $.  Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event. With this purpose in consideration, the $s$ value election is essential.
%necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
We propose the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
\begin{equation}
s_x = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
\label{eq:mapping2}
\end{equation}

@@

\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch. The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.  This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds in the horizontal direction. Finally, descriptors  $\mathbf{kp}$  are located at $ x_{kp} =  0.55 Fs.\gamma = 35 $ and $ y_{kp} = z(c) $ (see Eq.~\ref{eq:zerolevel}).   In this way the whole transient event is captured. 
Figure~\ref{fig:sampledescriptor2} shows a patch of a signal plot covering the complete amplitude (vertical direction) and the complete span of the signal event (horizontal direction). 
% esta figura no está comentada en el texto.
\begin{figure}[H]
\centering
\includegraphics[width=10cm]{patchgeometry.pdf}
\caption{The scale of local patch is selected in order to capture the whole signal information}
\label{fig:sampledescriptor2}
\end{figure}

%Hacemos algún comentario de k_a?
